---
title: "Lesson2"
output: 
   html_document:
     keep_md: true
---

<br>

**Defining function:**

```{r}
fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9
  return(temp_C)
}
```

**Using function:**
```{r}
fahrenheit_to_celsius(32)
fahrenheit_to_celsius(212)
```

<br>

**Automatic returns:**

In R, **it is not necessary to include the return statement.** R automatically returns whichever variable is on the last line of the body of the function. While in the learning phase, we will explicitly define the return statement.

<br>

**Creating another function**

```{r}
celsius_to_kelvin <- function(temp_C) {
  temp_K <- temp_C + 273.15
  return(temp_K)
}
```

**Composing functions**

```{r}
fahrenheit_to_kelvin <- function(temp_F) {
  temp_C <- fahrenheit_to_celsius(temp_F)
  temp_K <- celsius_to_kelvin(temp_C)
  return(temp_K)
}
```

**Nesting Functions**
    
This example showed the output of `fahrenheit_to_celsius` assigned to `temp_C`, which is then passed to `celsius_to_kelvin` to get the final result. It is also possible to perform this calculation in one line of code, by “nesting” one function inside another, like so:

```{r}
# freezing point of water in Fahrenheit
celsius_to_kelvin(fahrenheit_to_celsius(32.0))
```
<br>

**Exercise**

In the last lesson, we learned to combine elements into a vector using the c function, e.g. `x <- c("A", "B", "C")` creates a vector `x` with three elements. Furthermore, we can extend that vector again using c, e.g. `y <- c(x, "D")` creates a vector y with four elements. Write a function called highlight that takes two vectors as arguments, called content and wrapper, and returns a new vector that has the wrapper vector at the beginning and end of the content.

Resolution:

```{r}
#Defining function:
highlight <- function(content, wrapper) {
  return(c(wrapper, content, wrapper))
}

#Using function
best_practice <- c("Write", "programs", "for", "people", "not", "computers")
asterisk <- "***"  # R interprets a variable with a single value as a vector
                   # with one element.

highlight(best_practice, asterisk)
```

**Exercise**

If the variable v refers to a vector, then `v[1]` is the vector’s first element and `v[length(v)]` is its last (the function length returns the number of elements in a vector). Write a function called edges that returns a vector made up of just the first and last elements of its input.

Resolution:

```{r}
edges <- function(vec) {
  return(c(vec[1], vec[length(vec)]))
}

edges(c("Don't", "repeat", "yourself", "or", "others"))

```

**Named variables and the scope of variables**

Functions can accept arguments explicitly assigned to a variable name in the function call `functionName(variable = value)`, as well as arguments by order:

```{r}
mySum <- function(input_1, input_2 = 10) { # input_2 default value is 10
  output <- input_1 + input_2
  return(output)
}

mySum(input_1 = 1, 3) #Overwrites default value for input_2
mySum(3) #Does not overwrite input_2 default value
mySum(input_2 = 4) #Will raise error, since input_2 does not have default value
mySum(3,5)
```
**Testing, error handling and Documenting**

There are many ways of adding documentation/error handling to functions. We can add error handling using the `warning` and `stop` functions.

A common way to put documentation in software is to add comments like this:

```{r}
center <- function(data, midpoint) {
  # return a new vector containing the original data centered around the
  # midpoint.
  # Example: center(c(1, 2, 3), 0) => c(-1, 0, 1)
  new_data <- (data - mean(data)) + midpoint
  return(new_data)
}
```

However, formal documentation for R functions is written in separate `.Rd` using a markup language similar to LaTeX. You see the result of this documentation when you look at the help file for a given function, e.g. ?read.csv. The `roxygen2` package allows R coders to write documentation alongside the function code and then process it into the appropriate `.Rd` files. You will want to switch to this more formal method of writing documentation when you start writing more complicated R projects.

**Functions to Create Graphs**

Write a function called analyze that takes a filename as an argument and displays the three graphs produced in the previous lesson (average, min and max inflammation over time). analyze("data/inflammation-01.csv") should produce the graphs already shown, while analyze("data/inflammation-02.csv") should produce corresponding graphs for the second data set. Be sure to document your function with comments.

```{r}
analyze = function (filename) {
  data = read.csv(filename, header = FALSE)
  avg_data = apply(data, 2, mean)
  max_data = apply(data, 2, max)
  min_data = apply(data, 2, min)
  plot(avg_data)
  plot(max_data)
  plot(min_data)
}

analyze('~/Dropbox/repos/r_learn/swcarpentry/data/inflammation-01.csv')
```


**Rescaling**

Write a function rescale that takes a vector as input and returns a corresponding vector of values scaled to lie in the range 0 to 1. (If L and H are the lowest and highest values in the original vector, then the replacement for a value v should be (v-L) / (H-L).) Be sure to document your function with comments.

Test that your rescale function is working properly using min, max, and plot.

```{r}
rescale = function(vector) {
  max_value = max(vector)
  min_value = min(vector)
  return ((vector - min_value) / (max_value - min_value))
}

rescale(1:10)

```



**A Function with Default Argument Values**

Rewrite the rescale function so that it scales a vector to lie between 0 and 1 by default, but will allow the caller to specify lower and upper bounds if they want. Compare your implementation to your neighbor’s: Do your two implementations produce the same results when both are given the same input vector and parameters?


```{r}
rescale = function(vector, lower_limit = 0, upper_limit = 1) {
  max_value = max(vector)
  min_value = min(vector)
  return ((vector - min_value) / (max_value - min_value) * (upper_limit - lower_limit) + lower_limit)
}

rescale(1:10)
rescale(1:10, lower_limit = 10, upper_limit = 20)
```


**Key Points**

- Define a function using `name <- function(...args...) {...body...}`.
 
- Call a function using `name(...values...)`.
 
- R looks for variables in the current stack frame before looking for them at the top level.
 
- Use `help(thing)` to view help for something.
 
- Put comments at the beginning of functions to provide help for that function.
 
- Annotate your code!
 
- Specify default values for arguments when defining a function using `name = value` in the argument list.
 
- Arguments can be passed by matching based on name, by position, or by omitting them (in which case the default value is used).