---
title: "lesson4"
author: "Gabriel Alves Vieira"
date: "1/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lesson 4 - Making Choices

## Overview 

**Questions**

- How do I make choices using if and else statements?

- How do I compare values?

- How do I save my plots to a PDF file?

**Objectives**

- Save plot(s) in a PDF file.

- Write conditional statements with if and else.

- Correctly evaluate expressions containing && (‘and’) and || (‘or’).


## Saving plots to a file


Setting previously written functions:

```{r}
dir.create("plots")

analyze <- function(filename) {
  # Plots the average, min, and max inflammation over time.
  # Input is character string of a csv file.
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation)
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation)
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation)
}

analyze_all <- function(folder = "data", pattern) {
  # Runs the function analyze for each file in the given folder
  # that contains the given pattern.
  filenames <- list.files(path = folder, pattern = pattern, full.names = TRUE)
  for (f in filenames) {
    analyze(f)
  }
}
```


**Challenge**

We can update the analyze function so that it always saves the plots in a pdf. But that would make it more difficult to interactively test out new changes. It would be ideal if analyze would either save or not save the plots based on its input.

```{r}
analyze <- function(filename, out_pdf = F) {
  # Plots the average, min, and max inflammation over time.
  # Input is character string of a csv file.
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  max_day_inflammation <- apply(dat, 2, max)
  min_day_inflammation <- apply(dat, 2, min)
  if (out_pdf) {
    pdf_path <- paste("./plots/",tools::file_path_sans_ext(basename(filename)), ".pdf", sep = "") #generates filepath for that specific pdf file based on basename of original file without extension
    pdf(pdf_path) # redirects plots to pdf file 
    plot(min_day_inflammation)
    plot(avg_day_inflammation)
    plot(max_day_inflammation)
    dev.off() # stop R from redirecting plots to pdf file
  }
  else {
    plot(min_day_inflammation)
    plot(avg_day_inflammation)
    plot(max_day_inflammation)
  }
}

analyze_all <- function(folder = "data", pattern, out_pdf = F) {
  # Runs the function analyze for each file in the given folder
  # that contains the given pattern.
  filenames <- list.files(path = folder, pattern = pattern, full.names = TRUE)
  for (f in filenames) {
    if (out_pdf) {
     analyze(f, out_pdf = T) 
    }
    else {
      analyze(f)
    }
  }
}

analyze_all(pattern = "inflammation", out_pdf = T)
```

**Overwriting plots**

If you run `pdf` multiple times without running `dev.off`, you will save plots to the most recently opened file. However, you won’t be able to open the previous pdf files because the connections were not closed. In order to get out of this situation, you’ll need to run `dev.off` until all the pdf connections are closed. You can check your current status using the function `dev.cur`. If it says “pdf”, all your plots are being saved in the last pdf specified. If it says “null device” or “RStudioGD”, the plots will be visualized normally.

## Conditionals

`if, else if, else` -> Keywords

`&&, ||, !` -> AND, OR, NOT operators

`==, !=, <=, >=` -> comparison operators


**Choosing Plots Based on Data**

Write a function `plot_dist` that plots a boxplot if the length of the vector is greater than a specified threshold and a stripchart otherwise. To do this you’ll use the R functions `boxplot` and `stripchart`.


**Answer**

```{r}
plot_dist <- function(vec, threshold) {
  if (length(vec) > threshold) {
    boxplot(vec)
  } else {
    stripchart(vec)
  }
}
```



```{r}
dat <- read.csv("data/inflammation-01.csv", header = FALSE)
plot_dist(dat[, 10], threshold = 10)     # day (column) 10

plot_dist(dat[1:5, 10], threshold = 10)  # samples (rows) 1-5 on day (column) 10
```

**Histograms Instead**

One of your collaborators prefers to see the distributions of the larger vectors as a histogram instead of as a boxplot. In order to choose between a histogram and a boxplot we will edit the function `plot_dist` and add an additional argument `use_boxplot`. By default we will set `use_boxplot` to `TRUE` which will create a boxplot when the vector is longer than threshold. When `use_boxplot` is set to `FALSE`, `plot_dist` will instead plot a histogram for the larger vectors. As before, if the length of the vector is shorter than threshold, `plot_dist` will create a stripchart. A histogram is made with the `hist` command in R.


```{r}
plot_dist <- function(vec, threshold, use_boxplot = T) {
  if (length(vec) > threshold && use_boxplot) {
    boxplot(vec)
  } else if (length(vec) > threshold && !use_boxplot) {
    hist(vec)
  } else {
    stripchart(vec)
  }
}
```

```{r}
dat <- read.csv("data/inflammation-01.csv", header = FALSE)
plot_dist(dat[, 10], threshold = 10, use_boxplot = TRUE)   # day (column) 10 - create boxplot

plot_dist(dat[, 10], threshold = 10, use_boxplot = FALSE)  # day (column) 10 - create histogram

plot_dist(dat[1:5, 10], threshold = 10)                    # samples (rows) 1-5 on day (column) 10
```


**Find the Maximum Inflammation Score**

Find the file containing the patient with the highest average inflammation score. Print the file name, the patient number (row number) and the value of the maximum average inflammation score.


Tips:

- Use variables to store the maximum average and update it as you go through files and patients.
- You can use nested loops (one loop is inside the other) to go through the files as well as through the patients in each file (every row).

Complete the code below:

```{r}

filenames <- list.files(path = "data", pattern = "inflammation-[0-9]{2}.csv", full.names = TRUE)
filename_max <- "" # filename where the maximum average inflammation patient is found
patient_max <- 0 # index (row number) for this patient in this file
average_inf_max <- 0 # value of the average inflammation score for this patient
for (f in filenames) {
  dat <- read.csv(file = f, header = FALSE)
  dat.means <- apply(dat, 1, mean)
  for (patient_index in 1:length(dat.means)){
    patient_average_inf <- dat.means[patient_index]
    if (patient_average_inf > average_inf_max){ # If the patient average is higher than the current average_inf_max, update the variables
      average_inf_max <- patient_average_inf
      patient_max <- patient_index
      filename_max <- f
    }
  }
}
print(filename_max)
print(patient_max)
print(average_inf_max)
```

**Changing the Behavior of the Plot Command**

One of your collaborators asks if you can recreate the figures with lines instead of points. Find the relevant argument to plot by reading the documentation (?plot), update analyze, and then recreate all the figures with analyze_all.


```{r}
dir.create("plots")

analyze <- function(filename, output = NULL) {
  # Plots the average, min, and max inflammation over time.
  # Input:
  #    filename: character string of a csv file
  #    output: character string of pdf file for saving
  if (!is.null(output)) {
    pdf(output)
  }
  dat <- read.csv(file = filename, header = FALSE)
  avg_day_inflammation <- apply(dat, 2, mean)
  plot(avg_day_inflammation, type = "l")
  max_day_inflammation <- apply(dat, 2, max)
  plot(max_day_inflammation, type = "l")
  min_day_inflammation <- apply(dat, 2, min)
  plot(min_day_inflammation, type = "l")
  if (!is.null(output)) {
    dev.off()
  }
}

analyze_all <- function(pattern) {
  # Directory name containing the data
  data_dir <- "data"
  # Directory name for results
  results_dir <- "plots"
  # Runs the function analyze for each file in the current working directory
  # that contains the given pattern.
  filenames <- list.files(path = data_dir, pattern = pattern)
  for (f in filenames) {
    pdf_name <- file.path(results_dir, sub(".csv", ".pdf", f))
    analyze(file.path(data_dir, f), output = pdf_name)
  }
}

analyze_all("inflammation.*csv")
```


## Key Points

- Save a plot in a pdf file using `pdf("name.pdf")` and stop writing to the pdf file with `dev.off()`.
 
- Use `if (condition)` to start a conditional statement, `else if (condition)` to provide additional tests, and `else` to provide a default.
 
- The bodies of conditional statements must be surrounded by curly braces `{ }`.
 
- Use `==` to test for equality.
 
- `X && Y` is only true if both X and Y are true.
 
- `X || Y` is true if either X or Y, or both, are true.